---
title: "World Football R"
subtitle: "Scouting, shot maps & transfer values with worldfootballR"
categories: [R, Sport, ggplot2, worldfootballR]
date: 2026-04-25
---

## Overview

This analysis showcases the power of the `worldfootballR` package — a single R package that connects to **three major football data sources** to build a complete player analysis.

We focus on the top 4 players of the 2023 season : Kane, Mbappé, Haaland and Rodri.

| Source | Data | Function |
|------------------------|------------------------|------------------------|
| FBref | Scouting reports (per 90 stats) | `fb_player_scouting_report()` |
| Understat | Shot locations & xG | `understat_player_shots()` |
| Transfermarkt | Transfer & market values | `tm_player_transfer_history()` |

## Setup

```{r}
#| message: false
#| warning: false

library(worldfootballR)
library(tidyverse)
library(fmsb)
library(ggsoccer)
library(patchwork)
library(ggtext)
library(plotly)
```

------------------------------------------------------------------------

## Part 1 — Player profiles : Radar charts

Using FBref scouting reports, we compare 11 key metrics per 90 minutes across the Big 5 leagues.

```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show radar chart code"

data_player <- function(url) {

  player_data <- fb_player_scouting_report(
    url, "primary", time_pause = 3,
    league_comp_name = "Last 365 Days Men's Big 5 Leagues, UCL, UEL"
  )

  parameters <- c(
    "Goals", "xG: Expected Goals", "Pass Completion %", "Key Passes",
    "Assists", "Touches", "Successful Take-Ons", "Tackles",
    "GCA (Defensive Action)", "Blocks", "Clearances"
  )

  max_min <- data.frame(
    Goals = c(1, 0), `xG: Expected Goals` = c(1, 0),
    `Pass Completion %` = c(100, 0), Assists = c(1, 0),
    `Key Passes` = c(4, 0), `GCA (Defensive Action)` = c(0.05, 0),
    Tackles = c(40, 0), Blocks = c(1, 0), Clearances = c(2, 0),
    Touches = c(300, 0), `Successful Take-Ons` = c(10, 0)
  )

  study_player <<- player_data |>
    filter(Statistic %in% parameters) |>
    filter(StatGroup != "Standard") |>
    select(Statistic, Per90) |>
    pivot_wider(names_from = Statistic, values_from = Per90)

  name_player    <- paste0(player_data[1, "Player"], " (", player_data[1, "Versus"], ")")
  names(max_min) <- names(study_player)
  df             <- rbind(max_min, study_player)
  rownames(df)   <- c("max", "min", name_player)
  names(df)      <- c("Goals", "Exp. Goal", "Pass completion", "Assist",
                      "Key Pass", "GCA (Defensive)", "Tackles", "Blocks",
                      "Clearance", "Touches", "Successful Take-Ons")

  colors_border <- c(rgb(0.2, 0.5, 0.5, 0.9), rgb(0.8, 0.2, 0.5, 0.9))
  colors_in     <- c(rgb(0.2, 0.5, 0.5, 0.4), rgb(0.8, 0.2, 0.5, 0.4))

  radarchart(df, axistype = 1,
    pcol = colors_border, pfcol = colors_in, plwd = 4, plty = 1,
    cglcol = "grey", cglty = 1, axislabcol = "grey",
    caxislabels = seq(0, 1, 0.2), cglwd = 1, vlcex = 0.8
  )

  legend(x = -0.8, y = -1.2,
    legend = rownames(df[-c(1, 2), ]),
    bty = "n", pch = 10, col = colors_in,
    text.col = "grey", cex = 1, pt.cex = 3
  )
}

par(mfrow = c(2, 2), mar = c(1, 1, 1, 1), mgp = c(2, 1, 0), xpd = NA)
data_player("https://fbref.com/en/players/21a66f6a/Harry-Kane")
data_player("https://fbref.com/en/players/42fd9c7f/Kylian-Mbappe")
data_player("https://fbref.com/en/players/1f44ac21/Erling-Haaland")
data_player("https://fbref.com/en/players/6434f10d/Rodri")
mtext("Top 4 Players 2023", side = 3, line = -1.3,
      outer = TRUE, cex = 1.5, font = 4, col = "#0F6E56")
```

> **Key insight** : Rodri's radar looks completely different from the other three — dominant on defensive actions and passing, minimal on goals. This perfectly illustrates how a defensive midfielder compares to attackers.

------------------------------------------------------------------------

## Part 2 — Shot maps : Where do they score ?

Using Understat data, we plot the exact location of every goal scored in the 2023/24 season, with shot type (foot/head) as colour.

```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show shot map code"

mbappe  <- understat_player_shots("https://understat.com/player/3423") |>
           filter(season == 2023, result == "Goal") |>
           mutate(shotType = as.factor(shotType))

rodri   <- understat_player_shots("https://understat.com/player/2496") |>
           filter(season == 2023, result == "Goal") |>
           mutate(shotType = as.factor(shotType))

kane    <- understat_player_shots("https://understat.com/player/647") |>
           filter(season == 2023, result == "Goal") |>
           mutate(shotType = as.factor(shotType))

haaland <- understat_player_shots("https://understat.com/player/8260") |>
           filter(season == 2023, result == "Goal") |>
           mutate(shotType = as.factor(shotType))

shotType_levels <- c("Head", "LeftFoot", "OtherBodyPart", "RightFoot")
mbappe$shoTypeT  <- factor(mbappe$shotType,  levels = shotType_levels)
rodri$shoTypeT   <- factor(rodri$shotType,   levels = shotType_levels)
kane$shoTypeT    <- factor(kane$shotType,    levels = shotType_levels)
haaland$shoTypeT <- factor(haaland$shotType, levels = shotType_levels)

map_shot <- function(data, player_name) {
  data <- data |> mutate(x = X * 100, y = Y * 100)
  n    <- nrow(data)

  ggplot(data) +
    annotate_pitch(colour = "white", fill = "springgreen4", limits = FALSE) +
    geom_jitter(aes(x = x, y = y, colour = shoTypeT),
                width = 1, height = 1, size = 2) +
    theme_pitch() +
    theme(panel.background = element_rect(fill = "springgreen4")) +
    coord_flip(xlim = c(49, 101)) +
    scale_y_reverse() +
    scale_color_discrete(drop = FALSE) +
    ggtitle(player_name, paste0("Number of goals: ", n)) +
    theme(panel.border  = element_rect(color = "black", fill = NA, size = 3),
          plot.title    = element_text(hjust = 0.5))
}

plot1 <- map_shot(mbappe,  "Kylian Mbappé")  + guides(colour = "none")
plot2 <- map_shot(rodri,   "Rodri")          + guides(colour = "none")
plot3 <- map_shot(kane,    "Harry Kane")     + guides(colour = "none")
plot4 <- map_shot(haaland, "Erling Haaland")

plot1 + plot2 + plot3 + plot4 +
  plot_layout(guides = "collect") +
  plot_annotation(
    title    = "Goal positions — Top 4 players",
    subtitle = "Season 2023/2024",
    caption  = "Source: understat.com via worldfootballR"
  )
```

The interactive version below lets you hover over each goal to see the xG value, shot type, assist provider, and match details.

```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show interactive version"

map_shot_v <- function(data, player_name) {
  data <- data |>
    mutate(x = X * 100, y = Y * 100,
           xG2    = round(xG, 2),
           match_ = paste0(home_team, " - ", away_team))
  n <- nrow(data)

  p <- ggplot(data) +
    annotate_pitch(colour = "white", fill = "springgreen4", limits = FALSE) +
    geom_point(aes(x = x, y = y, label1 = xG2, label2 = shotType,
                   label3 = player_assisted, label4 = match_),
               size = 2, colour = "blue") +
    theme_pitch() +
    theme(panel.background = element_rect(fill = "springgreen4")) +
    coord_flip(xlim = c(49, 101)) +
    scale_y_reverse() +
    ggtitle(paste0(player_name, " — Goals 2023/2024"))

  ggplotly(p, tooltip = c("label1", "label2", "label3", "label4"))
}

map_shot_v(kane, "Harry Kane")
```

> **Key insight** : Haaland scores almost exclusively inside the box — a pure penalty-area striker. Mbappé shows a wider spread, reflecting his ability to create and finish from various positions.

------------------------------------------------------------------------

## Part 3 — Transfer market values

Using Transfermarkt data, we track how each player's market value has evolved over their career.

```{r}
#| eval: false
#| code-fold: true
#| code-summary: "Show transfer value code"

haaland <- tm_player_transfer_history("https://www.transfermarkt.com/erling-haaland/profil/spieler/418560")
mbappe  <- tm_player_transfer_history("https://www.transfermarkt.com/kylian-mbappe/profil/spieler/342229")
kane    <- tm_player_transfer_history("https://www.transfermarkt.com/harry-kane/profil/spieler/132098")
rodri   <- tm_player_transfer_history("https://www.transfermarkt.com/rodri/profil/spieler/357565")

kane_2    <- kane    |> filter(complete.cases(pick(transfer_date, market_value)))
rodri_2   <- rodri   |> filter(complete.cases(pick(transfer_date, market_value)))
haaland_2 <- haaland |> filter(complete.cases(pick(transfer_date, market_value)))
mbappe_2  <- mbappe  |> filter(complete.cases(pick(transfer_date, market_value)))

the_four <- bind_rows(haaland_2, mbappe_2, kane_2, rodri_2)

p <- ggplot(the_four, aes(x = transfer_date, y = market_value, color = player_name)) +
  geom_line() +
  geom_point(aes(label1 = team_to, label2 = market_value)) +
  labs(x = "Date", y = "Market Value",
       title  = "Top 4 — Transfer Market Value History",
       colour = "Player") 

ggplotly(p, tooltip = c("label1", "label2"))
```

> **Key insight** : Mbappé's value peaked earliest and highest — a reflection of his extraordinary rise at Monaco and PSG. Haaland's curve is the steepest, driven by his Bundesliga and Premier League performances.

------------------------------------------------------------------------

## Video walkthroughs

### Part 1 — Radar charts & player profiles

{{< video https://youtu.be/c4misumzbkI >}}

### Part 2 — Shot maps & transfer values

{{< video https://youtu.be/OTWw1kaPuFg >}}

## Conclusion

With just a few lines of R and `worldfootballR`, we connected to 3 different data sources and built a complete multi-dimensional player analysis :

-   **Who are they** as players ? → Radar charts
-   **How do they score** ? → Shot maps\
-   **What is their value** ? → Transfer history

## Source code

Full code available on [GitHub](https://github.com/FlipR-data/worldFootballR).
